\startcomponent c_4_interview_conclusion
\product prd_Chapter4
\project project_thesis

\subsection{Reflection Conclusion}

My findings fundamentally agree with and extend both Chaim Zins' work and the definitions of the {\em Oxford English Dictionary}. While more research is necessary, there is little doubt in my mind that I have answered the two questions of interest: 

\startitemize
\item I have discovered constructions of data that profoundly differ from each other and that agree with O, related research. 
\item Some evidence points to the existence of trading zones between different constructions of data. The use of the terms \quotation{raw data} and \quotation{derived data} in the interviews strongly suggests this. However, more research is necessary before there is any useful evidence for how trading zones form with constructions of data and what impact evaluative accents have on data and misunderstandings.
\item There is evidence for both Ackoff's and Tuomi's hierarchies of data, information, and knowledge. Ackoff's work has penetrated modern thinking, especially management thinking, quite deeply. At the same time, the more \quotation{modern} relativistic philosophies of science seem to correspond quite well with Tuomi's ontology. Far more evidence is needed to create suggested \quotation{trading zones} that can bridge these and other discovered conceptions of data.
\item Some indications hint that the developer’s understanding of data is a factor in database design, though the current evidence cannot create useful rules for database changes based on the recipient's current conception of data. More implementation research is needed here, as well. 
\item Both the interviews and the surveys discovered different realities of data, but not without failures. More research is necessary to explore the utility of my methodology and to refine it for different purposes. 
\stopitemize


\stopcomponent
